---Title: "Census Fuel Type Analysis"
Author: "John Cornwell"
Date: "December 21, 2017"
Output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pkgs <- list("plyr", "dplyr", "tidyr")
lapply(pkgs, library, character.only = TRUE) # load the required packages
table(CAFuel.County$NAME)
```

# Censusapi Fuel Type Analysis Request

We would want to know things like the following, but I’d happily take whatever info is actually available (so it doesn’t have to be as defined below):
a) share of CA hhlds that use alternative fuels (probably mostly propane or wood, but maybe diesel, fuel oil, or coal) as a primary fuel of some sort (most likely for space or water heating)
b) share of IOU customers in CA that use alternative fuels
c) which fuels for either a or b
d) any geographic pockets of high incidence
e) if available, how these numbers look for LI hhlds
f) anything we know about how many hhlds are entirely off the grid in CA

## Identify dataset of interest. 
### ACS 5 year vintage 2016
```{r}
# list API endpoints and find the one that corresponds to SAHIE.

APIList <- listCensusApis()
name: acs/acs5
vintage 2016
```

## Excellent - now we have found the dataset we need to find out what is in the dataset. We need to get a list of variables so we can find the one we want.

To do this we use the **listCensusMetadata** function. 
```{r}

acs5.2016_vars <- listCensusMetadata(name="acs/acs5", vintage = 2016, type = "variables")
acs5.2016_vars
acs5.2016_fuel_vars <- filter(acs5.2016_vars, grepl("FUEL",concept))
print(acs5.2016_fuel_vars)
acs5.2016_fuel_vars[, c(1, 2)] 
```
Fuel Vars:
STATE: STATE FIPS CODE
COUNTY:
TRACT:
BLKGRP:
B25040_001E	Estimate!!Total			
B25040_008E	Estimate!!Total!!Solar energy			
B25040_009E	Estimate!!Total!!Other fuel			
B25040_006E	Estimate!!Total!!Coal or coke			
B25040_007E	Estimate!!Total!!Wood			
B25040_004E	Estimate!!Total!!Electricity			
B25040_005E	Estimate!!Total!!Fuel oil, kerosene, etc.			
B25040_002E	Estimate!!Total!!Utility gas			
B25040_003E	Estimate!!Total!!Bottled, tank, or LP gas	
```{r}
# get heating fuel count for CA in 2016 5 year ACS.
CAFuel <- getCensus(name="acs/acs5", vintage = 2016,,
          vars=c("STATE", "B25040_001E","B25040_008E","B25040_009E", "B25040_006E",
                 "B25040_007E","B25040_004E","B25040_005E","B25040_002E","B25040_003E"), 
          region="state:06")
CAFuel.t <- as.data.frame(t(CAFuel))
```
County

```{r}
CAFuel.County <-  getCensus(name="acs/acs5", vintage = 2016,
                  vars=c("COUNTY", "NAME", "B25040_001E","B25040_008E","B25040_009E", "B25040_006E","B25040_007E","B25040_004E","B25040_005E","B25040_002E","B25040_003E"), 
                  region= "county:*", regionin =  "state:06")
CAFuel.County$AltFuel <- rowSums(CAFuel.County[c("B25040_006E","B25040_007E","B25040_008E","B25040_009E","B25040_005E", "B25040_003E")], na.rm = T)
CAFuel.County$PercAltF <- CAFuel.County$AltFuel/CAFuel.County$B25040_001E
CAFuel.County.t <- as.data.frame(t(CAFuel.County))
```

Tracts
```{r}
CAFuel.Tract <-  getCensus(name="acs/acs5", vintage = 2016,,
                  vars=c("GEO_ID","COUNTY", "NAME", "B25040_001E","B25040_008E","B25040_009E", "B25040_006E",
                  "B25040_007E","B25040_004E","B25040_005E","B25040_002E","B25040_003E"), 
                  region= "tract:*", regionin =  "state:06")
CAFuel.Tract.t <- as.data.frame(t(CAFuel.Tract))

```

Import and merge IOU geo flags from Athens Research data onto tracts
```{r}
IOU_GEO_TRACT <- read.csv('ATHENS_BG_ONLY_UTILITYFLAG.csv', stringsAsFactors = F)
IOU_GEO_TRACT <- unique(IOU_GEO_TRACT)
IOU_GEO_TRACT$value <- 1
IOU_GEO_TRACT$util <- as.factor(IOU_GEO_TRACT$util)
IOU_GEO_TRACT.wide <- spread(IOU_GEO_TRACT, util, value)

CAFuel.Tract$GEO_ID <- substring(CAFuel.Tract$GEO_ID, 8) 

CAFuel.Tract.IOUGEO <- left_join(CAFuel.Tract, IOU_GEO_TRACT.wide, "GEO_ID")

CAFuel.Tract.IOUGEO$IOU <- rowSums(CAFuel.Tract.IOUGEO[c("PGEEG", "PGEEO","PGEGO","SCEE","SCGG","SDGEE","SDGEG")], na.rm = T)
CAFuel.Tract.IOUGEO$isIOU <- ifelse(CAFuel.Tract.IOUGEO$IOU > 0,1,0)

write.csv(CAFuel.Tract.IOUGEO,"CAFuel.Tract.IOUGEO.csv")
write.csv(CAFuel.County,"counties.csv")


table(IOU_GEO_TRACT$util)
table(CAFuel.Tract.IOUGEO$isIOU)
```

LI by county

```{r}
CAPov.County <-  getCensus(name="acs/acs5", vintage = 2016,
                  vars=c("COUNTY", "NAME", "B17026_001E","B17026_002E","B17026_003E", "B17026_004E", "B17026_005E","B17026_006E","B17026_007E","B17026_008E","B17026_009E", "B17026_010E", "B17026_011E"), 
                  region= "county:*", regionin =  "state:06")
write.csv(CAPov.County,"CAPov.County.csv")

CAFuel.County.t <- as.data.frame(t(CAFuel.County))
```


## Map of alternative fuel penetration - county
```{r }
library(FSA)
library(ggplot2)
rename(d, c("beta"="two", "gamma"="three"))

CAFuel.County <- plyr::rename(CAFuel.County, c("NAME" = "subregion"))

CAFuel.County.map <- CAFuel.County %>% group_by(subregion) %>% summarise(mean=mean(PercAltF,na.rm=T),                                                                      sd=sd(PercAltF,na.rm=T))

CAFuel.County.map$subregion <- tolower(CAFuel.County.map$subregion)

ca2 <-map_data('county',"california")

ca2_join <- left_join(ca2, cntygroupformap2,by="subregion")
ca2_join <- unique(ca2_join, by = subregion)

ca_lookup <- ca2_join %>% distinct(subregion, .keep_all = T)

table(ca_lookup$Quartile)

## set directory
setwd("~/R/HEUS-Paper/")

##readOGR. This function, from the rgdal package, automatically extracts the information regarding the data.
ca <- readOGR(dsn = "California County Shape Files/County", layer = "CaliforniaCounty")
ca
ca_ordered <- ca[ order(ca[,c("NAME")])]
extract <- (ca$NAME)
extract
ca_f <-fortify(ca)
head(ca_f)
table(ca_f$id)
extract <- as.data.frame(extract)
names(extract)[names(extract) == 'extract'] <- 'subregion'
extract$subregion <- as.character(extract$subregion)
extract$subregion <- tolower(extract$subregion)
ca_lookup2<- left_join(extract,ca_lookup, by = 'subregion')
ca_lookup <- ca_lookup2

ca_lookup$id <- row.names(ca_lookup) # allocate an id variable to the sp data
ca_lookup$id <- as.numeric(ca_lookup$id)
ca_lookup$id <- ca_lookup$id-1
ca_lookup$id <- as.character(ca_lookup$id)
ca_f <- merge(x = ca_f, y = ca_lookup[ , c("id", "Quartile", "mean", "totaln", "ci95","subregion")], by = "id", all.x=TRUE) # join the data

table(ca_f$subregion,ca_f$Quartile)
centroid_ca_f <- ddply(ca_f, c("subregion"), summarise,
               mean.long = mean(long),
               mean.lat   = mean(lat))

ca_f <- merge(x = ca_f, y = centroid_ca_f, by = "subregion", all.x=TRUE) # join the data

ca_f$Quartile_1_2 <- NA
ca_f$Quartile_1_2 <- ifelse(ca_f$Quartile == '1', 1, NA)
ca_f$Quartile_1_2 <- ifelse(ca_f$Quartile == '2', 1, ca_f$Quartile_1_2)
ca_f$Quartile_1_2 <- ifelse(ca_f$Quartile == '3', 2, ca_f$Quartile_1_2)
ca_f$Quartile_1_2 <- ifelse(ca_f$Quartile == '4', 3, ca_f$Quartile_1_2)
ca_f$Quartile_1_2 <- as.factor(ca_f$Quartile_1_2)


map <- ggplot(ca_f, aes(long, lat, group = group, fill = Quartile)) + geom_polygon() + coord_equal() +
labs(fill = "County Efficiency Quartiles") + ggtitle("PG&E Service Territory Lighting By County Efficiency Quartiles")  + geom_path( data = ca_f , colour = "black")

map2 <- ggplot(ca_f, aes(long, lat, group = group, fill = Quartile_1_2)) + geom_polygon() + coord_equal() + labs(fill = "County \nEfficiency Quartiles")  + geom_path( data = ca_f , colour = "black") +  theme(plot.title = element_text(hjust = 0.5)) + scale_fill_manual(labels = c("Quartiles 1-2","Quartile 3","Quartile 4"), breaks = c("1", "2", "3"),values = c("grey80","#00BFC4","#C77CFF"))

table(ca_f$Quartile_1_2)

map2
scale_color_manual(breaks = c("1", "2", "3"), values = c("blue","purple","grey50")) +
cnty_id_tbl_caf <- table(ca_f$subregion, ca_f$id)
cnty_id_tbl_ca_lookup <- table(ca_lookup$subregion, ca_lookup$id)
cnty_id_tbl_ca_lookup <- table(ca_lookup$subregion, ca_lookup$id)

write.csv(cnty_id_tbl_ca_lookup,"cnty_id_tbl_ca_lookup.csv")
write.csv(cnty_id_tbl_caf,"cnty_id_tbl_caf.csv")
write.csv(cnty_id_tbl_caf,"cnty_id_tbl_cafmerge.csv")

+ geom_text(data=ca_f, aes(mean.long, mean.lat, label = subregion), size=1)

map2
map3
```
Blocks -- NOT DONE
```{r}
CAFuel.Blocks <-  getCensus(name="acs/acs5", vintage = 2016,
                  vars=c("BLKGRP"), 
                  region=  "county")
CAFuel.Tract.t <- as.data.frame(t(CAFuel.Tract))


IOU_GEO_TRACT_Counties <- IOU_GEO_TRACT.wide
IOU_GEO_TRACT_Counties$CntyID <- substring(IOU_GEO_TRACT_Counties$GEO_ID, 5,7)
IOU_GEO_TRACT_Counties <- IOU_GEO_TRACT_Counties[!duplicated(IOU_GEO_TRACT_Counties$CntyID), ]
write.csv(IOU_GEO_TRACT_Counties,"IOUCounty.csv")

```
